# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data

```{r}

Sys.setlocale("LC_ALL", "english")

activity <- read.csv("activity.csv")
activity$date <- as.Date(activity$date, "%Y-%M-%d")
```

## What is mean total number of steps taken per day?
```{r}
library(lattice)
#with(activity, hist(tapply(steps, date, sum, na.rm = TRUE), xlab = "total number of steps taken"))
with(activity, histogram(tapply(steps, date, sum, na.rm = TRUE), xlab = "total number of steps taken"))
print(m1 <- with(activity, mean(tapply(steps, date, sum, na.rm = TRUE))))
print(m2 <-with(activity, median(tapply(steps, date, sum, na.rm = TRUE))))
```
The mean value of total number of steps taken per day is `r m1`
The median value of total number of steps take per day is `r m2`


## What is the average daily activity pattern?
```{r}
#with(activity, plot(unique(interval), tapply(steps, interval, mean, na.rm = TRUE), type = "l", xlab = "", ylab = "Ave Steps/interval"))
xyplot(tapply(steps, interval, mean, na.rm = TRUE) ~ unique(interval), data = activity, type = "l", xlab = "", ylab = "Ave Steps/interval")
print(max_interval_ind <- with(activity, which.max(tapply(steps, interval, mean, na.rm = TRUE))))
print(max_interval <- with(activity, tapply(steps, interval, mean, na.rm = TRUE)[[max_interval_ind]]))
```
the 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps is in interval `r names(max_interval_ind)` and maxium average of steps by interval is `r as.numeric(max_interval)`
## Imputing missing values
```{r}
print(num_missing <- nrow(activity[is.na(activity$steps), ]))
```
The total number of missing values in the dataset is  `r num_missing`

Because it seems that the observations of certain days are all missing, it is preferrable to use median value of 5-minute interval to replace missing value
```{r}
m2 <- with(activity, tapply(steps, interval, median, na.rm = TRUE))
complete_activity <- activity
complete_activity$steps <- mapply(function(s, i) ifelse(is.na(s), m2[[as.character(i)]], s), activity$steps, activity$interval)
```
```{r}
with(complete_activity, histogram(tapply(steps, date, sum), xlab = "total number of steps taken"))
print(m1 <- with(complete_activity, mean(tapply(steps, date, sum, na.rm = TRUE))))
print(m2 <-with(complete_activity, median(tapply(steps, date, sum, na.rm = TRUE))))
```

## Are there differences in activity patterns between weekdays and weekends?
an additional factor variable w_day_flag is created 
```{r}
#Create a new factor variable in the dataset with two levels ¨C ¡°weekday¡± and ¡°weekend¡± #indicating whether a given date is a weekday or weekend day
complete_activity <- within(complete_activity, wday_flag <- factor(ifelse(weekdays(activity$date) %in% c("Saturday", "Sunday"), "weekend", "weekday")))
weekday_act <- as.data.frame(with(subset(complete_activity, wday_flag == "weekday"), tapply(steps, interval, mean)))
colnames(weekday_act) <- "steps"
weekday_act$w_day_flag <- "weekday"
weekday_act$interval <- rownames(weekday_act)

weekend_act <- as.data.frame(with(subset(complete_activity, wday_flag == "weekend"), tapply(steps, interval, mean)))
colnames(weekend_act) <- "steps"
weekend_act$w_day_flag <- "weekend"
weekend_act$interval <- rownames(weekend_act)

act_sum <- rbind(weekday_act, weekend_act)
act_sum$w_day_flag <- as.factor(act_sum$w_day_flag)
act_sum$interval <- as.numeric(act_sum$interval)
xyplot(steps ~ interval | w_day_flag, data = act_sum, type = "l", layout = c(1, 2), ylab = "number of steps")
```
